Symmetric Peer-to-Peer Applications

### 2022-08-24 tpd-22.tex
...

In this section, we describe the architecture of our middleware and also the
API in which developers have to conform with.

@@

The main job of the middleware is to deal with the uncertainties of the
network in order to preserve the symmetric behavior across peers.
More specifically, the middleware ensures that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric:
%
\begin{enumerate}
\item \textbf{Event Deadlines:}
Due to the inherent latency of networks, peers generate events with a
deliberate delay so that they reach other peers in time to be applied in sync
in a common deadline.
%We first assume a predetermined delay for the sake of simplicity.
However, a peer that is locally ahead of time may receive an event that should
have been applied in the past, event considering the delay.
In this case, the middleware automatically rolls back the peer, applies the
event at the correct time, and then fast forwards the peer, re-applying all
remaining events up to the real time.
%
\item \textbf{Time Synchronization:}
Since peers run in different machines, their timelines will inevitably be out
of sync because:
    (a) applications are launched locally at different times, and
    (b) internal clocks may diverge over time.
The middleware assumes the maximum time among all peers the be the
``correct real time'', so that other peers must fast forward to catch up with
it.
Therefore, in order to determine the maximum time and synchronize the clocks,
the middleware proceeds as follows:
    (a) peers broadcast \code{SYNC} events periodically with their local
        times, and
    (b) if a peer is behind a received \code{SYNC}, it advances its frames
        proportionally faster to catch up in time.
%
\item \textbf{Peer Churn:}
Considering that we target peer-to-peer networks, it is important that the
middleware can handle arrival and departure of peers seamlessly.
TODO
\end{enumerate}

@@

The middleware expects to take full control of the application execution in
order to control its timeline, and also to communicate events in the network.
For this reason, the basic programming API is to call a \code{loop} function
passing a set of callbacks, which the middleware uses to communicate with the
application:

@@

Under the hood, the middleware relies on the library \emph{SDL}%
\footnote{SDL: \url{www.libsdl.org}}
to deal with the network and input events, such as timers, mouse clicks and
key presses.

@@

{\footnotesize
\begin{verbatim}
void cb_ini (int ini) {
    if (ini) {
        // create the SDL window
        // open the arrow PNG images
        // create peer links (peer-id, ip, port)
        p2p_link(2, "192.168.1.2", 5000);
    } else {
        // destroy the SDL window
        // destroy the arrow PNG images
        // destroy peer links
        p2p_unlink(2);
    }
}
\end{verbatim}
}

@@

{\footnotesize
\begin{verbatim}
void cb_sim (p2p_evt evt) {
    switch (evt.id) {
        case P2P_EVT_INIT:  // resets position and speed
            G.x = G.y = G.dx = G.dy = 0;
            break;
        case P2P_EVT_TICK:  // increases position
            G.x += G.dx * VEL;
            G.y += G.dy * VEL;
            break;
        case P2P_EVT_KEY:   // modifies speed
            switch (evt.pay.i1) {
                case SDLK_UP:
                    G.dx = 0;
                    G.dy = -1;
                    break;
                <...> // similar to DOWN/LEFT/RIGHT/SPACE
            }
            break;
    }
}
\end{verbatim}
}

@@

{\footnotesize
\begin{verbatim}
void cb_eff (void) {
    // clears the SDL window
    // redraws the rectangle at the current position
    SDL_Rect r = { G.x, G.y, 20, 20 };
    SDL_DrawRect(&r);
}
\end{verbatim}
}

>>>>

In this section, we describe the architecture of our middleware and also the
API in which developers have to conform with.
The middleware is written in C and relies on the library \emph{SDL}%
\footnote{SDL: \url{www.libsdl.org}}
to deal with the network and input events, such as timers, mouse clicks and
key presses.
The full source code%
\footnote{TODO}
is under 500 lines of code.

@@

The main job of the middleware is to deal with the uncertainties of the
network in order to preserve the symmetric behavior across peers.
More specifically, the middleware ensures that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric:
%
\begin{enumerate}
\item \textbf{Event Deadlines:}
Due to the inherent latency of networks, peers generate events with a
deliberate delay so that they reach other peers in time to be applied in sync
in a common deadline.
%We first assume a predetermined delay for the sake of simplicity.
However, a peer that is locally ahead of time may receive an event that should
have been applied in the past, event considering the delay.
In this case, the middleware automatically rolls back the peer, applies the
event at the correct time, and then fast forwards the peer, re-applying all
remaining events up to the real time.
%
\item \textbf{Time Synchronization:}
Since peers run in different machines, their timelines will inevitably be out
of sync because:
    (a) applications are launched locally at different times, and
    (b) internal clocks may diverge over time.
The middleware assumes the maximum time among all peers the be the
``correct real time'', so that other peers must fast forward to catch up with
it.
Therefore, in order to determine the maximum time and synchronize the clocks,
the middleware proceeds as follows:
    (a) peers broadcast \code{SYNC} events periodically with their local
        times, and
    (b) if a peer is behind a received \code{SYNC}, it advances its frames
        proportionally faster to catch up in time.
%
\item \textbf{Peer Churn:}
Considering that we target peer-to-peer networks, it is important that the
middleware can handle arrival and departure of peers seamlessly.
When a peer leaves the network, nothing needs to be done.
The middleware just needs to ensure that a peer that joins the network receives
all events in order, which is trivial since they all hold an absolute
timestamp.
\end{enumerate}

@@

The middleware expects to take full control of the application execution in
order to control its timeline and disseminate events in the network.
For this reason, the basic programming API is to call a \code{loop} function
passing a set of callbacks, which the middleware uses to communicate with the
application:

@@

@@

{\footnotesize
\begin{verbatim}
void cb_ini (int ini) {
    if (ini) {
        <...> // create the SDL window
        <...> // open the arrow PNG images
        // create peer links (peer-id, ip, port)
        p2p_link(2, "192.168.1.2", 5000);
    } else {
        <...> // destroy the SDL window
        <...> // destroy the arrow PNG images
        // destroy peer links
        p2p_unlink(2);
    }
}
\end{verbatim}
}

@@

{\footnotesize
\begin{verbatim}
void cb_sim (p2p_evt evt) {
    switch (evt.id) {
        case P2P_EVT_INIT: // resets pos and speed
            G.x = G.y = G.dx = G.dy = 0;
            break;
        case P2P_EVT_TICK:  // increases position
            G.x += G.dx * VEL;
            G.y += G.dy * VEL;
            break;
        case P2P_EVT_KEY:  // modifies speed
            switch (evt.pay.i1) {
                case SDLK_UP:
                    G.dx = 0;
                    G.dy = -1;
                    break;
                <...> // same for other keys
            }
            break;
    }
}
\end{verbatim}
}

@@

{\footnotesize
\begin{verbatim}
void cb_eff (void) {
    <...> // clears the SDL window
    // redraws the rectangle at the current position
    SDL_Rect r = { G.x, G.y, 20, 20 };
    SDL_DrawRect(&r);
}
\end{verbatim}
}
